---
title: "Homework 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)

```

## Problem 1

Read the Mr. Trashwheel dataset

```{r}
trashwheel_df = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "Mr. Trash Wheel", 
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read the percipitation data for 2018 and 2017

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2018 Precipitation", 
    skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
  
precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2017 Precipitation", 
    skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
  
```

Now combine annual precipitation

```{r}

month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")


```

This dataset contains information from the Mr. Trashwheel trash colector in Baltimore, MD. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including guitars. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. 

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.


## Problem 2

Read, clean, and tidy the NYC transit dataset

```{r}
transit_df = 
  read_csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, vending, entry, ada) %>% 
  mutate(
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)
  ) %>% 
  mutate(
    entry = as.logical(recode(entry, "YES" = "TRUE", "NO" = "FALSE"))
  )

transit_df_tidy =
  transit_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "route",
    names_prefix = "route",
    values_to = "route_name"
  ) %>% 
  drop_na(route_name) %>% 
  distinct(line, station_name, route, route_name, .keep_all = TRUE) %>% 
  relocate(line, station_name, route, route_name)

```

The NYC transit dataset contains geographical information about stations and lines in NYC and the routes for train lines associated with them. There are `r n_distinct(transit_df_tidy$line)` lines, specifically (`r paste(unique(transit_df_tidy$line), collapse = ", ")`). The initial table was not tidy. The second dataframe I created, transit_df_tidy, is tidy because its in the longer format.

* The total number of stations are `r nrow(transit_df_tidy %>% distinct(line, route_name))`.
* The total number of stations that are ADA compliant are `r nrow(transit_df_tidy %>% filter(ada == TRUE) %>% distinct(line, route_name))`.
* The proportion of station entrances / exits without vending allow entrance is `r round(nrow(transit_df_tidy %>% filter(vending == "NO", entry == TRUE) %>% distinct(line, route_name)) / nrow(transit_df_tidy %>% filter(vending == "NO") %>% distinct(line, route_name)),2)`.

* The A train is served by `r nrow(transit_df_tidy %>% filter(route_name == "A") %>% distinct(line, route_name))` stations.
* Of those stations, `r nrow(transit_df_tidy %>% filter(route_name == "A") %>% filter(ada == TRUE) %>% distinct(line, route_name))` are ADA compliant. 


## Problem 3

Read and load pols-month, snp.csv, and unemployment datasets

```{r}
pols_month_df =
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    month = as.integer(month)
  ) %>% 
  left_join(month_df, by = "month") %>% 
  relocate(month_name, .after = year) %>% 
  mutate(
    president = if_else(prez_gop == 1, "GOP", "DEM")
  ) %>% 
  select(-month, -day, -prez_gop, -prez_dem) %>% 
  arrange(year, month_name)

snp_df =
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/")%>% 
  mutate(
    month = as.integer(month)
  ) %>% 
  left_join(month_df, by = "month") %>% 
  relocate(month_name, .after = year) %>% 
  select(-month, -day) %>% 
  arrange(year, month_name)

month_df =
  tibble(
    month_name = month.name,
    month = tolower(month.abb)
  )

unemployment_df =
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names()

unemployment_df_tidy =
  unemployment_df %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment_rate"
  ) %>% 
  left_join(month_df, by = "month") %>% 
  select(-month) %>% 
  arrange(year, month_name)

  
```

